---
title: "IBS 519 Homework 3 Revision Report"
author: "Firstname Lastname"
date: "2022-09-07"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# knitr options to set up
# show code in report
knitr::opts_chunk$set(echo = TRUE)
# if errors, keep knitting
knitr::opts_chunk$set(error = TRUE)
# hide messages and warnings
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

# load printr package for nicer
# table output formatting
library(printr)

# load packages needed
library(readr)
library(dplyr)
library(ggplot2)
library(purrr)

# load abalone dataset
abalone <- read_csv("abalone.csv")
```

## Data cleaning - dimensional measurements

Previously in class, we reviewed the 3 dimensional measurements of the abalones. We know from the documentation provided at the data source at [https://archive.ics.uci.edu/ml/datasets/abalone](https://archive.ics.uci.edu/ml/datasets/abalone), that length should be the longest dimension and should be > height and > diameter.

This code will list the id's of the abalones with any issues - including any with illogical values (such as height = 0).

### Summary Statistics of dimensional measurements

In the summary stats reported below, you'll notice there are some abalones with height = 0.

```{r}
# code from in class examples
abalone %>%
  select(length, diameter, height) %>%
  summary()
```

### Identify the abalones with height = 0

```{r}
abalone %>%
  filter(height == 0) %>%
  select(id, length, diameter, height)
```

### Identify the abalones with height > length

```{r}
abalone %>%
  filter(height > length) %>%
  select(id, length, diameter, height)
```

### Identify the abalones with diameter > length

```{r}
abalone %>%
  filter(diameter > length) %>%
  select(id, length, diameter, height)
```

### Visualize the abalone with height > length

From the plot below, we can see any abalones above the reference line that have height > length and using color for each sex, we can also note which sex has this illogical measurement.

The abalone with height > length is a F, female.

There is also a M, male, with a large height relative to its length (even though its height is < its length), and it's height is quite a bit larger than the other abalone heights.

```{r}
ggplot(abalone, aes(length, height)) +
  geom_point(aes(color = sex)) +
  geom_abline(intercept = 0,
              slope = 1,
              color = "blue")
```

### Other possible outliers for height

The histogram below shows 2 abalones with larger heights than the rest. One of these abalones is the abalone with height > length. The other abalone is just large with values that are possible but will have higher leverage in future analyses.

```{r}
ggplot(abalone, aes(height)) + 
  geom_histogram(aes(y=..density..), 
                 binwidth=.02,
                 colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666")
```

### Show the abalones with the largest heights

The table below shows the abalones with the top 10 largest heights.

```{r}
# add slice_head() function to specify number
# of rows to show
abalone %>% 
  select(id, sex, length, diameter, height) %>%
  arrange(desc(height)) %>%
  slice_head(n = 10)
```


## Data cleaning - weight measurements

Add your code from Homework 3 below to show how to check and identify any of the abalones with illogical weight measurements. As noted in the documentation, wholeWeight should be the largest weight, so check to see if any abalones have:

* shuckedWeight > wholeWeight
* visceraWeight > wholeWeight
* shellWeight > wholeWeight

### TASK 1: Identify the abalones with shuckedWeight > wholeWeight

```{r}
# PUT YOUR CODE HERE
```

### TASK 2: Identify the abalones with visceraWeight > wholeWeight

```{r}
# PUT YOUR CODE HERE
```

### TASK 3: Identify the abalones with shellWeight > wholeWeight

```{r}
# PUT YOUR CODE HERE
```

### TASK 4: Show the abalones that have shuckedWeight > wholeWeight

Make a plot below similar to the one above that shows the abalones with shuckedWeight > wholeWeight. Any abalones above the reference line have illogical weights. And the colors are also shown by sex to identify the sex of the abalones with the illogical shuckedWeight.

```{r}
# PUT YOUR CODE HERE
```

### TASK 5: Write description here

WRITE A FEW SENTENCES HERE ON WHAT YOU SEE IN THE PLOT FOR WHICH SEX ABALONES HAVE THESE OUTLIERS. WHY DO YOU THINK THESE ABALONES HAD ILLOGICAL VALUES?

## TASK 6: Summary statistics of cleaned dataset

```{r}
# PUT YOUR CODE HERE TO GET CLEANED DATASET
abalone_cleaned <- abalone %>%
  filter(height > 0) %>%         
  filter(diameter <= length) %>% 
  filter(height <= length) %>%
  filter(___) %>%
  filter(___) %>%
  filter(___)
```

Now that we've cleaned up the dataset, let's get the summary statistics for the various measurements on these abalones by sex.

### Summary statistics of the 3 Dimensional Measurements by Sex

```{r}
# This approach uses the purrr package
library(purrr)
abalone_cleaned %>%
  select(sex, length, diameter, height) %>%
  split(.$sex) %>%
  map(summary)
```

### TASK 7: Summary statistics of the 4 Weight Measurements by Sex

```{r}
# This approach uses the purrr package
library(purrr)
abalone_cleaned %>%
  select(sex, ___, ___, ___, ___) %>%
  split(.$sex) %>%
  map(summary)
```
